### 2021-12-03 Draft Final Project1.Rmd
Add files via upload

---
title: "Final Project Data 205"
author: "Tiffany King"
date: "11/18/2021"
output: html_document
---

# Loading Libraries
```{r}
library('rvest')
library(tidyverse)
library(psych)
library(readr)
library(dplyr)
library(lubridate)
library(ggplot2)
library(plotly)
library(plyr)
library(highcharter)
library(RColorBrewer)
```
# Setting working directory
```{r}
setwd("/Users/tiffanyking/Desktop/DATA 205/DATA205/DATA205")
```

```{r}
WKridership<- read_csv("2019 Ridership weekday average.csv")
crimes<-read_csv("crime.csv")
``` 
# Examine the data
```{r}
str(WKridership)
head(WKridership)
```
## create dataframe from CSV
```{r}
year<- c("2010","2011","2012","2013","2014","2015","2016","2017","2018","2019","2020","2021")
averider<- c(705563,737313,721225,706591,693705,675077,621079,613665,605910,626269,176602,125557)
dailyentriesride<-data.frame(year,averider)
print(dailyentriesride)
```
# Barplot Average Daily Entries 
```{r}
options(scipen = 999)
ggplot(data=dailyentriesride, aes(x=year, y=averider)) +
  geom_bar(stat="identity", fill="red")+
  geom_text(aes(label=averider), vjust=1.6, color="white", size=3.5)+
   xlab("Year") +
  ylab("Passengers") +
  ggtitle("Average Daily Entries By Train") +
  ylim(0,800000)
  theme_minimal()
```
##Barplot Average Daily Entries by Bus

```{r}
year<- c("2010","2011","2012","2013","2014","2015","2016","2017","2018","2019","2020","2021")
averider<- c(705563,737313,721225,706591,693705,675077,621079,613665,605910,626269,176602,125557)
dailyentriesride<-data.frame(year,averider)
print(dailyentriesride)
```

## Getting the WMATA Data
```{r}
library(tidyverse)
library(pdftools)
library(httr)
```

```{r}
get_wmata_data <- function(pdflink) {
  wmata_data <- pdf_text(pdflink) %>%
    str_split("\n")

  df = ""

  for (i in 1:length(wmata_data)) {

    for (line in wmata_data[[i]][9:(length(wmata_data[[i]])-2)]) {
      if (line == "" | is.na(line))
      {
        next
      } else {
        line <- str_squish(line)

        if (substr(line,21,27) %in% c("SUMMONS"))
        {
          stub <- str_split_fixed(line, " - ", 2)
          state <- substr(stub[1], nchar(stub[1])-1, nchar(stub[1]))
          stub_1 <- substr(stub[1],1, nchar(stub[1])-3)
          stub_2 <- stub[2]
          line <- paste(stub_1, stub_2, "\n")
          line <- paste(substr(line, 1, 36), state, " ", substr(line, 37, nchar(line)))
          prev_line <- line
          #print(line)
        }

        if (substr(line,21,27) %in% c("ARREST ", "REPORT "))
        {
          line <- paste(substr(line,1,26), "        ", substr(line,27,nchar(line)))
          stub <- str_split_fixed(line, " - ", 2)
          state <- substr(stub[1], nchar(stub[1])-1, nchar(stub[1]))
          stub_1 <- substr(stub[1],1, nchar(stub[1])-3)
          stub_2 <- stub[2]
          line <- paste(stub_1, stub_2, "\n")
          line <- paste(substr(line, 1, 36), state, " ", substr(line, 37, nchar(line)))
          prev_line <- line
          #print(line)
        }

        if (substr(line,3,3) !="/")
        {
          offense_stub <- str_split_fixed(line, " - ", 2)

          if (offense_stub[2]  != "")
          {
            line <- paste(stub_1, substr(line,6,nchar(line)))
            line <- paste(substr(line, 1, 36), state, " ", substr(line, 37, nchar(line)), "\n")
          }else{
            df <- substr(df, 1, nchar(df)-nchar(prev_line))
            line <- paste(substr(prev_line,1,nchar(prev_line)-2), offense_stub[1], "\n")
          }
        }

        df <- paste(df, line)
        print(line)
      }
    }
  }
  return(read_fwf(df, 
                  col_types=cols('c','c','c','c','c'),
                  fwf_cols(date=12, time=9, type=17, state=4, desc=NA)))
}
#fwf_cols(date=12, time=9, type=17, state=4, desc=NA),
# Now define what the links to the pdf files should look like and create a
# way to iterate through them to get multiple files at one go.

url <- "https://www.wmata.com/about/transit-police/upload/Monthly-Blotter-"
time_periods <- c("October-2021","September-2021", "August-2021", "July-2021", "June-2021","May-2021","April-2021","March-2021","February-2021","January-2021")

url <- "https://www.wmata.com/about/transit-police/upload/Monthly-Blotter-"

time_periods <- c("December-2020","November-2020","October-2020","September-2020", "August-2020", "July-2020", "June-2020","May-2020","April-2020","March-2020","February-2020","January-2020")

df_prev = ""

for (time_period in time_periods) {
  pdflink <- paste0(url, time_period, ".pdf")
  df_curr <- get_wmata_data(pdflink)

  if (df_prev == "")
  {
    df_prev <- df_curr
    df_wmata <- df_prev
  } else {
    df_wmata <- bind_rows(df_wmata, df_curr)
  }
}

# Check to make sure the data starts with start month and ends with end month.
df_wmata
tail(df_wmata)

```

# Create Plot
```{r}
df.groupby('year').type.value_counts().unstack(0).plot.barh()

```

### 2021-12-03 Draft Final Project1.Rmd
Add files via upload

@@
 url <- "https://www.wmata.com/about/transit-police/upload/Monthly-Blotter-"
-time_periods <- c("October-2021","September-2021", "August-2021", "July-2021", "June-2021","May-2021","April-2021","March-2021","February-2021","January-2021")
+time_periods <- c("October-2021","September-2021", "August-2021", "July-2021", "June-2021","May-2021","April-2021","March-2021","February-2021","January-2021","December-2020","November-2020","October-2020","September-2020", "August-2020")
 
-url <- "https://www.wmata.com/about/transit-police/upload/Monthly-Blotter-"
-
-time_periods <- c("December-2020","November-2020","October-2020","September-2020", "August-2020", "July-2020", "June-2020","May-2020","April-2020","March-2020","February-2020","January-2020")
+url <-"https://www.wmata.com/about/transit-police/upload/June-2020-Blotter-Report"
 
@@
 
-# Create Plot
+## Clean the WMATA data with NA's
 ```{r}
-df.groupby('year').type.value_counts().unstack(0).plot.barh()
 
 ```
+
+# Create Groupby Year & State (test)
+```{r}
+df_state<- df_wmata %>%
+group_by(year) %>%
+select(DC,MD,VA)
+summarize(sum = sum(state), count = n()) %>%
+arrange(year, state)
+
+df_state
+```
+
+# Create Plot on the Crimes Committed by State
+```{r}
+
+```
+# Creat a Groupby type & count (tester)
+```{r}
+df_wmata %>%                         # Count rows by group
+  group_by(type) %>% 
+  summarize(n = n())
+```
+
+```{r}
+df_type <-df_wmata
+  group_by(type)
+df_type
+```
+
+```{r}
+df_type<- df_wmata %>%
+ group_by(type) %>%
+    summarize(count=n())
+
+```
+
+# Create Plot on Type of Crimes by Type
+```{r}
+ggplot(df_wmata, aes(x=type, fill=type )) + 
+  geom_bar( ) +
+  scale_fill_hue(c = 40) +
+  theme(legend.position="none")
+```
+
+# Create Plot on 
+```{r}
+
+```
+
+#Census Data 
+```{r}
+library(httr)
+library(jsonlite)
+library(tidyverse)
+
+# Import json data from Census API
+data <- GET("https://api.census.gov/data/2019/acs/acs1/?get=NAME,B01001_001E,B01002_001E,C22001_001E,B25064_001E&for=county:*&in=state:24")
+data <- content(data, as="text")
+
+# Load into a data frame
+df <- fromJSON(data)
+class(df)
+head(df)
+
+# Change to tibble
+df <- as_tibble(df)
+head(df)
+
+# Make first row column header names
+names(df) <- c(df[1,])
+head(df)
+
+# Keep all but first row (which had the column header names)
+df <- df[-1,]
+head(df)
+
+# Rename columns as needed
+names(df) <- c('NAME',
+               'B01001_001E' = 'Population',
+               'B01002_001E' = 'Median_Age',
+               'C22001_001E' = 'Household_Poverty',
+               'B25064_001E' = 'Median_Gross_Rent',
+               'state',
+               'county')
+head(df)
+
+```

### 2021-12-05 Draft Final Project1.Rmd
Add files via upload

@@
   theme_minimal()
-```
-##Barplot Average Daily Entries by Bus
-
-```{r}
-year<- c("2010","2011","2012","2013","2014","2015","2016","2017","2018","2019","2020","2021")
-averider<- c(705563,737313,721225,706591,693705,675077,621079,613665,605910,626269,176602,125557)
-dailyentriesride<-data.frame(year,averider)
-print(dailyentriesride)
 ```
@@
 url <- "https://www.wmata.com/about/transit-police/upload/Monthly-Blotter-"
-time_periods <- c("October-2021","September-2021", "August-2021", "July-2021", "June-2021","May-2021","April-2021","March-2021","February-2021","January-2021","December-2020","November-2020","October-2020","September-2020", "August-2020")
+time_periods <- c("October-2021","September-2021", "August-2021", "July-2021", "June-2021","May-2021","April-2021","March-2021","February-2021","January-2021","December-2020","November-2020","October-2020","September-2020","August-2020","May-2020","April-2020","March-2020","February-2020","January-2020","December-2019","November-2019","October-2019","June-2019","November-2018")
 
-url <-"https://www.wmata.com/about/transit-police/upload/June-2020-Blotter-Report"
+url_exception_june_2020<-"https://www.wmata.com/about/transit-police/upload/June-2020-Blotter-Report.pdf"
+df_exception_June_2020 <- get_wmata_data(url_exception_june_2020)
+
+url_exception_july_2020<-"https://www.wmata.com/about/transit-police/upload/July-2020-Blotter.pdf"
+df_exception_July_2020 <- get_wmata_data(url_exception_july_2020)
+
+url_exception_july_2019<-"https://www.wmata.com/about/transit-police/upload/July-2019-Blotter.pdf"
+df_exception_July_2019 <- get_wmata_data(url_exception_July_2019)
+
+url_exception_may_2019<-"https://www.wmata.com/about/transit-police/upload/May-2019-Blotter-Report.pdf"
+df_exception_May_2019 <- get_wmata_data(url_exception_May_2019)
+
+url_exception_april_2019<-"https://www.wmata.com/about/transit-police/upload/Blotter-Report-April-2019.pdf"
+df_exception_April_2019 <- get_wmata_data(url_exception_April_2019)
+
+url_exception_march_2019<-"https://www.wmata.com/about/transit-police/upload/2019-March-Blotter-Report.pdf"
+df_exception_March_2019 <- get_wmata_data(url_exception_March_2019)
+
+url_exception_february_2019<-"https://www.wmata.com/about/transit-police/upload/February-2019-Monthly-Blotter-Report.pdf"
+df_exception_February_2019 <- get_wmata_data(url_exception_February_2019)
+
+url_exception_january_2019<-"https://www.wmata.com/about/transit-police/upload/January-2019-Monthly-Blotter-Report.pdf"
+df_exception_January_2019 <- get_wmata_data(url_exception_January_2019)
+
+url_exception_december_2018<-"https://www.wmata.com/about/transit-police/upload/Monthly-Blotter-2018-December.pdf"
+df_exception_December_2018 <- get_wmata_data(url_exception_December_2018)
+
+url_exception_october_2018<-"https://www.wmata.com/about/transit-police/upload/October-2018-Monthly-Crime-Blotter.pdf"
+df_exception_October_2018 <- get_wmata_data(url_exception_October_2018)
+
+url_exception_september_2018<-"https://www.wmata.com/about/transit-police/upload/September-2018-Blotter.pdf"
+df_exception_September_2018 <- get_wmata_data(url_exception_September_2018)
+
+url_exception_august_2018<-"https://www.wmata.com/about/transit-police/upload/August-2018-Monthly-Blotter.pdf"
+df_exception_August_2018 <- get_wmata_data(url_exception_August_2018)
 
@@
   } else {
-    df_wmata <- bind_rows(df_wmata, df_curr)
+    df_wmata <- bind_rows(df_wmata, df_curr,url_exception_june_2020,url_exception_june_2020,url_exception_march_2019,url_exception_february_2019,url_exception_january_2019,url_exception_december_2018,url_exception_october_2018,url_exception_september_2018,url_exception_august_2018)
   }
@@
 tail(df_wmata)
+
+```
+
+```{r}
 

### 2021-12-07 Draft Final Project1.Rmd
Add files via upload

@@
 url <- "https://www.wmata.com/about/transit-police/upload/Monthly-Blotter-"
-time_periods <- c("October-2021","September-2021", "August-2021", "July-2021", "June-2021","May-2021","April-2021","March-2021","February-2021","January-2021","December-2020","November-2020","October-2020","September-2020","August-2020","May-2020","April-2020","March-2020","February-2020","January-2020","December-2019","November-2019","October-2019","June-2019","November-2018")
+time_periods <- c("October-2021","September-2021", "August-2021", "July-2021", "June-2021","May-2021","April-2021","March-2021","February-2021","January-2021","December-2020","November-2020","October-2020","September-2020","August-2020","May-2020","April-2020","March-2020","February-2020","January-2020","December-2019","November-2019","October-2019","June-2019","November-2018","March-2018","February-2018","January-2018","October-2017","September-2017")
 
@@
 
+url_exception_september_2019<-"https://www.wmata.com/about/transit-police/upload/September-2019-Monthly-Blotter.pdf"
+df_exception_September_2019 <- get_wmata_data(url_exception_september_2019)
+
+url_exception_august_2019<-"https://www.wmata.com/about/transit-police/upload/August-2019-Monthly-Blotter.pdf"
+df_exception_August_2019 <- get_wmata_data(url_exception_august_2019)
+
 url_exception_july_2019<-"https://www.wmata.com/about/transit-police/upload/July-2019-Blotter.pdf"
-df_exception_July_2019 <- get_wmata_data(url_exception_July_2019)
+df_exception_July_2019 <- get_wmata_data(url_exception_july_2019)
 
 url_exception_may_2019<-"https://www.wmata.com/about/transit-police/upload/May-2019-Blotter-Report.pdf"
-df_exception_May_2019 <- get_wmata_data(url_exception_May_2019)
+df_exception_May_2019 <- get_wmata_data(url_exception_may_2019)
 
 url_exception_april_2019<-"https://www.wmata.com/about/transit-police/upload/Blotter-Report-April-2019.pdf"
-df_exception_April_2019 <- get_wmata_data(url_exception_April_2019)
+df_exception_April_2019 <- get_wmata_data(url_exception_april_2019)
 
 url_exception_march_2019<-"https://www.wmata.com/about/transit-police/upload/2019-March-Blotter-Report.pdf"
-df_exception_March_2019 <- get_wmata_data(url_exception_March_2019)
+df_exception_March_2019 <- get_wmata_data(url_exception_march_2019)
 
 url_exception_february_2019<-"https://www.wmata.com/about/transit-police/upload/February-2019-Monthly-Blotter-Report.pdf"
-df_exception_February_2019 <- get_wmata_data(url_exception_February_2019)
+df_exception_February_2019 <- get_wmata_data(url_exception_february_2019)
 
 url_exception_january_2019<-"https://www.wmata.com/about/transit-police/upload/January-2019-Monthly-Blotter-Report.pdf"
-df_exception_January_2019 <- get_wmata_data(url_exception_January_2019)
+df_exception_January_2019 <- get_wmata_data(url_exception_january_2019)
 
 url_exception_december_2018<-"https://www.wmata.com/about/transit-police/upload/Monthly-Blotter-2018-December.pdf"
-df_exception_December_2018 <- get_wmata_data(url_exception_December_2018)
-
-url_exception_october_2018<-"https://www.wmata.com/about/transit-police/upload/October-2018-Monthly-Crime-Blotter.pdf"
-df_exception_October_2018 <- get_wmata_data(url_exception_October_2018)
+df_exception_December_2018 <- get_wmata_data(url_exception_december_2018)
 
 url_exception_september_2018<-"https://www.wmata.com/about/transit-police/upload/September-2018-Blotter.pdf"
-df_exception_September_2018 <- get_wmata_data(url_exception_September_2018)
+df_exception_September_2018 <- get_wmata_data(url_exception_september_2018)
 
 url_exception_august_2018<-"https://www.wmata.com/about/transit-police/upload/August-2018-Monthly-Blotter.pdf"
-df_exception_August_2018 <- get_wmata_data(url_exception_August_2018)
+df_exception_August_2018 <- get_wmata_data(url_exception_august_2018)
+
+url_exception_july_2018<-"https://www.wmata.com/about/transit-police/upload/July-2018-Blotter.pdf"
+df_exception_July_2018 <- get_wmata_data(url_exception_july_2018)
+
+url_exception_june_2018<-"https://www.wmata.com/about/transit-police/upload/June-2016-Blotter.pdf"
+df_exception_June_2018 <- get_wmata_data(url_exception_june_2018)
+
+url_exception_may_2018<-"https://www.wmata.com/about/transit-police/upload/May-Monthly-blotter-2018.pdf"
+df_exception_May_2018 <- get_wmata_data(url_exception_may_2018)
+
+url_exception_april_2018<-"https://www.wmata.com/about/transit-police/upload/April-Monthly-blotter.pdf"
+df_exception_April_2018 <- get_wmata_data(url_exception_april_2018)
+
+url_exception_december_2017<-"https://www.wmata.com/about/transit-police/upload/December-Monthly-Blotter.pdf"
+df_exception_December_2017 <- get_wmata_data(url_exception_december_2017)
+
+url_exception_november_2017<-"https://www.wmata.com/about/transit-police/upload/November-Monthly-Blotter.pdf"
+df_exception_November_2017 <- get_wmata_data(url_exception_november_2017)
+
+url_exception_august_2017<-"https://www.wmata.com/about/transit-police/upload/Monthly-Blotter-Report-August-2017.pdf"
+df_exception_August_2017 <- get_wmata_data(url_exception_august_2017)
+
+url_exception_july_2017<-"https://www.wmata.com/about/transit-police/upload/Monthly-Blotter-Report-July-2017.pdf"
+df_exception_July_2017 <- get_wmata_data(url_exception_july_2017)
+
+url_exception_june_2017<-"https://www.wmata.com/about/transit-police/upload/June-2017-blotter.pdf"
+df_exception_June_2017 <- get_wmata_data(url_exception_june_2017)
+
+url_exception_april_2017<-"https://www.wmata.com/about/transit-police/upload/April-2017-monthly-blotter.pdf"
+df_exception_April_2017 <-get_wmata_data(url_exception_april_2017)
+
+url_exception_march_2017<-"https://www.wmata.com/about/transit-police/upload/March-2017-monthly-blotter.pdf"
+df_exception_March_2017 <-get_wmata_data(url_exception_march_2017)
+
+url_exception_february_2017<-"https://www.wmata.com/about/transit-police/upload/January-2017-monthly-blotter.pdf"
+df_exception_February_2017 <-get_wmata_data(url_exception_february_2017)
+
+url_exception_january_2017<-"https://www.wmata.com/about/transit-police/upload/Monthly-Blotter_February-2017.pdf"
+df_exception_January_2017 <-get_wmata_data(url_exception_january_2017)
 
@@
   } else {
-    df_wmata <- bind_rows(df_wmata, df_curr,url_exception_june_2020,url_exception_june_2020,url_exception_march_2019,url_exception_february_2019,url_exception_january_2019,url_exception_december_2018,url_exception_october_2018,url_exception_september_2018,url_exception_august_2018)
+    df_wmata <- bind_rows(df_wmata,df_curr,df_exception_June_2020,df_exception_July_2020,df_exception_September_2019,df_exception_August_2019,df_exception_March_2019,df_exception_February_2019,df_exception_January_2019,df_exception_December_2018,df_exception_September_2018,df_exception_August_2018,df_exception_July_2018, df_exception_June_2018,df_exception_May_2018,df_exception_April_2018,df_exception_December_2017,df_exception_November_2017,df_exception_August_2017,df_exception_July_2017,df_exception_June_2017,df_exception_April_2017,df_exception_March_2017,df_exception_February_2017,df_exception_January_2017)
+
   }
@@
 ```
+## View df_wmata Data 
+```{r}
+view(df_wmata)
+```
+
+## Summary
+```{r}
+summary(df_wmata)
+```
+
+## Get rid of NAs
+```{r}
+WMATA <-na.omit(df_wmata)
+```
+
+## 
+```{r}
+tuition$tuition$name[grepl('State University of New York', tuition$name)] <- 
+```
+
+## Two links that will not load & has an error
+```{r}
+
+url_exception_october_2018<-"https://www.wmata.com/about/transit-police/upload/October-2018-Monthly-Crime-Blotter.pdf"
+df_exception_October_2018 <- get_wmata_data(url_exception_october_2018)
+
+url_exception_may_2017<-"https://www.wmata.com/about/transit-police/upload/May-2017-Blotter.pdf"
+df_exception_May_2017 <- get_wmata_data(url_exception_may_2017)
+
+```
 
 ```{r}
-
+,url_exception_june_2020,url_exception_july_2020,url_exception_september_2019,url_exception_august_2019,url_exception_july_2019,url_exception_march_2019,url_exception_february_2019,url_exception_january_2019,url_exception_december_2018,url_exception_september_2018,url_exception_august_2018,url_exception_july_2018,url_exception_june_2018,url_exception_may_2018,url_exception_april_2018,url_exception_december_2017,url_exception_november_2017,url_exception_june_2017,url_exception_april_2017,url_exception_march_2017,url_exception_january_2017
 ```
@@
 ```
+
+## Remove Crimes Data 
+```{r}
+na.omit(crimes)
+```
+
+```{r}
+view(crimes)
+```
+
+```{r}
+names(crimes)[names(crimes) == 'Crime Name1'] <- 'Offenses'
+```
+
+```{r}
+P1<-crimes %>%                        
+  group_by(type) %>% 
+  select("")
+  summarize(n = n())
+P1
+```
+
+```{r}
+ggplot(crimes, aes(x=Offenses, fill=Offenses)) + 
+  geom_bar( ) +
+  scale_fill_hue(c = 40) +
+  theme(legend.position="none")
+```

### 2021-12-07 Draft Final Project1.Rmd
Add files via upload

@@
 
-## 
+## Rename variables (State)
 ```{r}
-tuition$tuition$name[grepl('State University of New York', tuition$name)] <- 
+WMATA$state[grepl('PG',WMATA$state)]<- 'MD'
+WMATA$state[grepl('MC',WMATA$state)]<- 'MD'
+WMATA$state[grepl('RL',WMATA$state)]<- 'VA'
+WMATA$state[grepl('EX',WMATA$state)]<- 'VA'
+```
+
+## Create a New Column and Add Months and years to WMATA DataSet 
+```{r}
+mutate(date=as.Date(date,format="%Y"))
+```
+
+## Groupby Time Series( Define Time (Morning, DayTime, Night)
+
+```{r}
+#Create a time series 
+```
+
+## Plot 1
+```{r}
+p1<-WMATA  %>%
+  filter(state %in% c("DC","MD","VA")) %>%
+  ggplot(aes(x=state, fill= state))+
+  geom_bar()+
+  labs(title="Total of Crimes by State/District",x="State/District", y="Total")+
+  scale_fill_brewer(palette = "Dark2") + theme_minimal()
+p1 + theme_minimal()
+```
+
+## Plot Types of Crime
+```{r}
+p2<-WMATA  %>%
+  filter(type %in% c("ARREST","REPORT","SUMMONS/CITATION")) %>%
+  ggplot(aes(x=type, fill= type))+
+  geom_bar()+
+  labs(title="Classification of Type of Crime/ Reporting",x="Type", y="Total")+
+  scale_fill_brewer(palette = "Blues") + theme_minimal()
+p2 + theme_minimal()
 ```
@@
 
-```
-
-```{r}
-,url_exception_june_2020,url_exception_july_2020,url_exception_september_2019,url_exception_august_2019,url_exception_july_2019,url_exception_march_2019,url_exception_february_2019,url_exception_january_2019,url_exception_december_2018,url_exception_september_2018,url_exception_august_2018,url_exception_july_2018,url_exception_june_2018,url_exception_may_2018,url_exception_april_2018,url_exception_december_2017,url_exception_november_2017,url_exception_june_2017,url_exception_april_2017,url_exception_march_2017,url_exception_january_2017
-```
-
-## Clean the WMATA data with NA's
-```{r}
-
-```
-
-# Create Groupby Year & State (test)
-```{r}
-df_state<- df_wmata %>%
-group_by(year) %>%
-select(DC,MD,VA)
-summarize(sum = sum(state), count = n()) %>%
-arrange(year, state)
-
-df_state
 ```

### 2021-12-08 Draft Final Project1.Rmd
Add files via upload

@@
 
-## Create a New Column and Add Months and years to WMATA DataSet 
+## Separate year from date
 ```{r}
-mutate(date=as.Date(date,format="%Y"))
+
 ```
@@
 ```{r}
-#Create a time series 
+#Create a time series plot
 ```
@@
 ```
+## Facet_wrap on the types of reports throughout the years
+```{r}
+ggplot(data = WMATA,
+       mapping = aes(x = date, y = n, color = type)) +
+  geom_line() +
+  facet_grid(rows = vars(type), cols =  vars(genus))
+```
+
+## Create a plot on
+```{r}
+
+```
 
@@
 df_exception_May_2017 <- get_wmata_data(url_exception_may_2017)
-
-```
-
-# Create Plot on the Crimes Committed by State
-```{r}
-
-```
-# Creat a Groupby type & count (tester)
-```{r}
-df_wmata %>%                         # Count rows by group
-  group_by(type) %>% 
-  summarize(n = n())
-```
-
-```{r}
-df_type <-df_wmata
-  group_by(type)
-df_type
-```
-
-```{r}
-df_type<- df_wmata %>%
- group_by(type) %>%
-    summarize(count=n())
-
-```
-
-# Create Plot on Type of Crimes by Type
-```{r}
-ggplot(df_wmata, aes(x=type, fill=type )) + 
-  geom_bar( ) +
-  scale_fill_hue(c = 40) +
-  theme(legend.position="none")
-```
-
-# Create Plot on 
-```{r}
 
@@
 
+## Create on plot on Income & Population(Linear Regression/ Scatterplot with Population)
+```{r}
+
+```
+
 ## Remove Crimes Data 
 ```{r}
-na.omit(crimes)
+Moco_crimes<-na.omit(crimes)
 ```
 
+## Plot on Classifcation of Crime
 ```{r}
-view(crimes)
+c1<-crimes  %>%
+  filter(`Crime Name1` %in% c("Crime Against Person ","Crime Against Property","Crime Against Society")) %>%
+  ggplot(aes(x=`Crime Name1`, fill=`Crime Name1`))+
+  geom_bar()+
+  labs(title="Classification of Crime",x="", y="Total")+
+  scale_fill_brewer(palette = "Accent") + theme_minimal()
+c1 + theme_minimal()
 ```
 
+## 
 ```{r}
-names(crimes)[names(crimes) == 'Crime Name1'] <- 'Offenses'
+
 ```
-
-```{r}
-P1<-crimes %>%                        
-  group_by(type) %>% 
-  select("")
-  summarize(n = n())
-P1
-```
-
-```{r}
-ggplot(crimes, aes(x=Offenses, fill=Offenses)) + 
-  geom_bar( ) +
-  scale_fill_hue(c = 40) +
-  theme(legend.position="none")
-```

### 2021-12-12 Draft Final Project1.Rmd
Add files via upload

--- 2021-12-07 Draft Final Project1.Rmd
@@
 time_periods <- c("October-2021","September-2021", "August-2021", "July-2021", "June-2021","May-2021","April-2021","March-2021","February-2021","January-2021","December-2020","November-2020","October-2020","September-2020","August-2020","May-2020","April-2020","March-2020","February-2020","January-2020","December-2019","November-2019","October-2019","June-2019","November-2018","March-2018","February-2018","January-2018","October-2017","September-2017")
+```
 
+```{r}
 url_exception_june_2020<-"https://www.wmata.com/about/transit-police/upload/June-2020-Blotter-Report.pdf"
@@
 ```
+
 ## View df_wmata Data 
@@
 
-## Create a New Column and Add Months and years to WMATA DataSet 
+## Separate year from date
 ```{r}
-mutate(date=as.Date(date,format="%Y"))
+WMATA$date <- as.Date(WMATA$date, format="%m/%d/%Y")
+head(WMATA)      
 ```
 
-## Groupby Time Series( Define Time (Morning, DayTime, Night)
+```{r}
+WMATA$year<-as.numeric(format(WMATA$date,"%Y"))
+head(WMATA)
+```
+
+## Groupby Time Series( Define Time (Morning, Afternoon ,Evening, Night)
 
 ```{r}
-#Create a time series 
+#Create a time series plot
+WMATA$hour <- as.numeric(gsub(":.*$", "", WMATA$time))
+WMATA$cat <- with(WMATA,  ifelse(hour >= 5 & hour<=11, "morning",
+                            ifelse(hour>11 & hour<=16, "evening", "night")))
+WMATA
 ```
@@
 
+##Prepare the Data
+```{r}
+DMV<-WMATA %>%
+  select(type,year)%>%
+  filter(year >=2017  & type %in% c("ARREST","REPORT","SUMMONS/CITATION"))%>%
+  group_by(year)
+```
+
+```{r}
+head(DMV)
+```
+
+## Getting the WMATA Geocodes
+```{r}
+library(dplyr)
+library(stringr)
+library(tidygeocoder)
+
+# Extract the street address from desc using Regex.  Extract crime description also.
+df_wmata2 <- df_wmata %>% mutate(street_addr=str_remove(desc, "[[:upper:]]{2,}.+")) %>%
+  mutate(crime_desc=str_extract(desc, "[[:upper:]]{2,}.+"))
+df_wmata2 <- df_wmata2 %>% mutate(full_addr=paste(street_addr,state))
+
+#Filter on just Maryland addresses
+df_md <- df_wmata2 %>% filter(state=="MD")
+
+#Apply the geocode function to get lat/lon values
+coordinates <- df_md %>% geocode(full_addr)
+```
+
+```{r}
+P5<- ggplot(DMV, aes(y =year, fill = factor(type))) +
+  geom_boxplot() +
+  facet_wrap(~type, scale = "free") +
+  scale_y_continuous(name = "Average Life Expectancy") +
+  scale_x_discrete(labels = abbreviate, name = "Continent")
+
+P5
+```
+
+```{r}
+ggplot(DMV, aes(x=type)) + geom_histogram()
+```
+
+```{r}
+ggplot(diamonds, aes(x=price, y=color, fill=color)) +
+```
+
+```{r}
+p <- ggplot(DMV, aes(year, type)) + 
+  geom_abline() +
+  geom_jitter(width = 0.1, height = 0.1) 
+p + facet_wrap()
+```
+
+## Facet_wrap on the types of reports throughout the years
+```{r}
+ggplot(data = WMATA, aes(year, n)) +
+  geom_line(color = "steelblue", size = 1) +
+  geom_point(color = "steelblue") + 
+  labs(title = "New Marvel characters by alignment",
+       subtitle = "(limited to characters with more than 100 appearances)",
+       y = "Count of new characters", x = "") + 
+  facet_grid(~type==c("DC","MD","VA"))
+```
+
 ## Two links that will not load & has an error
@@
 
+## Create on plot on Income & Population(Linear Regression/ Scatterplot with Population)
+```{r}
+
+```
+
 ## Remove Crimes Data 
 ```{r}
-na.omit(crimes)
+Moco_crimes<-na.omit(crimes)
 ```
 
+## Plot on Classifcation of Crime
 ```{r}
-view(crimes)
+c1<-crimes  %>%
+  filter(`Crime Name1` %in% c("Crime Against Person","Crime Against Property","Crime Against Society")) %>%
+  ggplot(aes(x=`Crime Name1`, fill=`Crime Name1`))+
+  geom_bar()+
+  labs(title="Classification of Crime",x="", y="Total")+
+  scale_fill_brewer(palette = "Accent") + theme_minimal()
+c1 + theme_minimal()
 ```
 
+## 
 ```{r}
-names(crimes)[names(crimes) == 'Crime Name1'] <- 'Offenses'
+
 ```
-
-```{r}
-P1<-crimes %>%                        
-  group_by(type) %>% 
-  select("")
-  summarize(n = n())
-P1
-```
-
-```{r}
-ggplot(crimes, aes(x=Offenses, fill=Offenses)) + 
-  geom_bar( ) +
-  scale_fill_hue(c = 40) +
-  theme(legend.position="none")
-```

### 2021-12-14 Draft Final Project1.Rmd
Add files via upload

@@
 WMATA$hour <- as.numeric(gsub(":.*$", "", WMATA$time))
-WMATA$cat <- with(WMATA,  ifelse(hour >= 5 & hour<=11, "morning",
-                            ifelse(hour>11 & hour<=16, "evening", "night")))
+WMATA$timeofday <- ifelse(WMATA$hour >= 5 & WMATA$hour<=11, "morning",
+                            ifelse(WMATA$hour > 11 & WMATA$hour<=16, "afternoon",
+                            ifelse(WMATA$hour >16 & WMATA$hour <=19, "evening", "night"))) 
 WMATA
@@
   filter(state %in% c("DC","MD","VA")) %>%
-  ggplot(aes(x=state, fill= state))+
+  ggplot(aes(x=year, fill= state))+
   geom_bar()+
   labs(title="Total of Crimes by State/District",x="State/District", y="Total")+
-  scale_fill_brewer(palette = "Dark2") + theme_minimal()
+  scale_fill_brewer(palette = "BuPu") + theme_minimal()
 p1 + theme_minimal()
@@
 
+## Preparing the data (time)
+```{r}
+d3<- WMATA%>%
+  group_by(year, timeofday) %>%
+  dplyr::summarize(Freq=n())
+```
+
+##Fix Na's 
+```{r}
+d4 <-na.omit(d3)
+```
+
+## Facet
+```{r}
+ ggplot(data = d4,
+        mapping = aes(x = year, y = Freq)) +
+     geom_bar(stat="identity") +
+     facet_wrap(vars(timeofday)) +
+     theme_minimal()
+```
+
 ##Prepare the Data
@@
 
+## using count 
+```{r}
+WMATA %>%
+  group_by(year,timeofday) %>%
+  summarise(n = n()) %>%
+  mutate(Freq = n/sum(n))
+```
+
+## Time Series
+
+```{r}
+sp <- ggplot(WMATA, aes(x=year, )) + geom_point(shape=1)
+sp + facet_grid(. ~ cat)
+```
+##
+```{r}
+hp <- ggplot(tips, aes(x=total_bill)) + geom_histogram(binwidth=2,colour="white")
+```
+
 ## Getting the WMATA Geocodes
@@
 # Extract the street address from desc using Regex.  Extract crime description also.
-df_wmata2 <- df_wmata %>% mutate(street_addr=str_remove(desc, "[[:upper:]]{2,}.+")) %>%
+df_wmata2 <- WMATA %>% mutate(street_addr=str_remove(desc, "[[:upper:]]{2,}.+")) %>%
   mutate(crime_desc=str_extract(desc, "[[:upper:]]{2,}.+"))
@@
 ggplot(DMV, aes(x=type)) + geom_histogram()
-```
-
-```{r}
-ggplot(diamonds, aes(x=price, y=color, fill=color)) +
 ```
@@
 
+```{r}
+library(dplyr)
+library(httr)
+#library(jsonlite)
+library(stringr)
+library(tidygeocoder)
+
+# Extract the street address from desc using Regex.  Extract crime description also.
+df_wmata2 <- WMATA %>% mutate(street_addr=str_remove(desc, "[[:upper:]]{2,}.+")) %>%
+  mutate(crime_desc=str_extract(desc, "[[:upper:]]{2,}.+"))
+df_wmata2 <- df_wmata2 %>% mutate(full_addr=paste(street_addr,state))
+
+#Filter on just Maryland addresses
+df_md <- df_wmata2 %>% filter(state=="MD")
+
+#Apply the geocode function to get lat/lon values
+coordinates <- df_md %>% geocode(full_addr)
+
+coordinates$census_tract <- 0
+row_count <- nrow(coordinates)
+```
+
 ## Create on plot on Income & Population(Linear Regression/ Scatterplot with Population)
@@
 
-## 
+## f
 ```{r}

### 2021-12-15 Draft Final Project1.Rmd
Add files via upload

@@
 WMATA$year<-as.numeric(format(WMATA$date,"%Y"))
+WMATA$month<-as.numeric(format(WMATA$date,"%m"))
 head(WMATA)
@@
 WMATA$hour <- as.numeric(gsub(":.*$", "", WMATA$time))
-WMATA$timeofday <- ifelse(WMATA$hour >= 5 & WMATA$hour<=11, "morning",
-                            ifelse(WMATA$hour > 11 & WMATA$hour<=16, "afternoon",
-                            ifelse(WMATA$hour >16 & WMATA$hour <=19, "evening", "night"))) 
+WMATA$timeofday <- ifelse(WMATA$hour >= 5 & WMATA$hour<=11, "Morning",
+                            ifelse(WMATA$hour > 11 & WMATA$hour<=16, "Afternoon",
+                            ifelse(WMATA$hour >16 & WMATA$hour <=19, "Evening", "Night"))) 
 WMATA
@@
   labs(title="Classification of Type of Crime/ Reporting",x="Type", y="Total")+
-  scale_fill_brewer(palette = "Blues") + theme_minimal()
+  scale_fill_brewer(palette = "") + theme_minimal()
 p2 + theme_minimal()
@@
  ggplot(data = d4,
-        mapping = aes(x = year, y = Freq)) +
+        mapping = aes(x = year, y = Freq, fill=timeofday)) +
      geom_bar(stat="identity") +
      facet_wrap(vars(timeofday)) +
+      scale_fill_brewer(palette = "Spectral") +
      theme_minimal()
+
 ```
@@
   select(type,year)%>%
+  group_by(year, type)%>%
   filter(year >=2017  & type %in% c("ARREST","REPORT","SUMMONS/CITATION"))%>%
-  group_by(year)
+  dplyr::summarize(Freq=n())
 ```
@@
 ```
+```{r}
+ggplot(data = DMV,
+      mapping = aes(x = year, y = Freq)) +
+     geom_line(stat="identity") +
+     facet_wrap(vars(type)) +
+     scale_fill_brewer(palette = "Spectral") +
+     theme_minimal()
+```
 
-## using count 
+## Months & Count
 ```{r}
-WMATA %>%
-  group_by(year,timeofday) %>%
-  summarise(n = n()) %>%
-  mutate(Freq = n/sum(n))
+DF6<- WMATA%>%
+  group_by(month) %>%
+  dplyr::summarize(count=n())
+DF6
+```
+
+```{r}
+df7<-na.omit(DF6)
 ```
@@
 ## Time Series 
-
 ```{r}
-sp <- ggplot(WMATA, aes(x=year, )) + geom_point(shape=1)
-sp + facet_grid(. ~ cat)
-```
-##
-```{r}
-hp <- ggplot(tips, aes(x=total_bill)) + geom_histogram(binwidth=2,colour="white")
+P6<- ggplot(df7, aes(x=month, fill=month)) + 
+  geom_bar(stat="identity", color="black", 
+           position=position_dodge()) +
+  geom_errorbar(aes(ymin=len-sd, ymax=len+sd), width=.2,
+                 position=position_dodge(.9))
 ```
@@
 library(dplyr)
+library(httr)
+#library(jsonlite)
 library(stringr)
@@
 df_md <- df_wmata2 %>% filter(state=="MD")
+```
 
+```{r}
 #Apply the geocode function to get lat/lon values
-coordinates <- df_md %>% geocode(full_addr)
+coordinates <- df_md %>% geocode(full_addr, method="arcgis")
+
+coordinates$census_tract <- 0
+row_count <- nrow(coordinates)
+
+# Get FIPS code from FCC API based on lat/lon
+fcc_url = "https://geo.fcc.gov/api/census/block/find?"
+headers = "'Content-Type':'application/json'"
+unique_record = 0
+
+for (row in 1:row_count) {
+  if (!is.na(coordinates$lat[row]) & coordinates$time[row] != unique_record) {
+
+    full_url <- paste0(fcc_url, "latitude=", coordinates$lat[row],
+                       "&longitude=", coordinates$long[row], "&censusYear=2010&format=json")
+    fcc <- GET(full_url)
+    fcc_parsed <- content(fcc, as="parsed")
+    try(coordinates$census_tract[row] <- fcc_parsed$Block$FIPS, TRUE)
+
+    print(full_url)
+    print(coordinates$census_tract[row])
+    unique_record = coordinates$time[row]
+  }
+  Sys.sleep(1)
+}
+
+census_info <- GET("https://api.census.gov/data/2019/acs/acs5?get=NAME,B01003_001E,B02001_002E,B02001_003E&for=block%20group:*&in=state:24&in=county:031&in=tract:*")
+df_census <- fromJSON(content(census_info, as="text"))
+
+df_census
 ```
